---
title: 'Session 4 - Data manipulation II: Long and wide data, joins, strings and factors'
author: "Jeppe N Samuelsen"
date: "24 2 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1:
Primary keys uniquely identifies observations in its own table.
Foreign keys uniquely identifies observations in another table.
Surrogate keys are keys that are added to the table because it lacks some kind of primary key.

Primary keys and forein keys typically form relations that are 1-to-many.

Q2:
Now assuming that we are working with two tables, x and y:

The inner_join() matches pairs of observations from mutiple tables (eg. x and y) when their keys are equal.

The left_join() keeps all observations from unique keys in x and matches with the observations of y that has a corresponding key in x.

The right_join() keeps all observations from unique keys in y and matches with the observations of x that has a corresponding key in y.

The full_join() keeps all observations in x and y.

Semi- and anti-join works on observations, not variables (keys):

The semi_join() keeps the observations of x that have a match in y.

The anti_join() drops all observations in x that have a match in y.

Q3:
inner_join() would be:
tribble(
~key, ~val_x, val_y,
1, "x1", "y1",
2, "x2", "y2"
)

left_join() would be:
tribble(
~key, ~val_x, val_y,
1, "x1", "y1",
2, "x2", "y2"
3, "x3", "NA"
)

right_join() would be:
tribble(
~key, ~val_x, val_y,
1, "x1", "y1",
2, "x2", "y2"
4, "NA", "y3"
)

full_join() would be:
tribble(
~key, ~val_x, val_y,
1, "x1", "y1",
2, "x2", "y2"
3, "x3", "NA"
4, "NA", "y3"
)

semi_join() would be:
tribble(
~key, ~val_x,
1, "x1",
2, "x2"
)

anti_join() would be:
tribble(
~key, ~val_x,
3, "x3"
)

Q4:
Creating function, which returns a random dna string of length n:
```{r}
library(tidyverse)

dna_string <- function() {
  my_dna <- str_c(sample(x = c("A", "T", "C", "G"), size = 100, replace = TRUE), collapse = "")
  }
```
Finding fraction of adenine in the dna:
```{r}
str_count(string = my_dna, pattern = "A")
```

25/100 = 1/4 of the bases in the string are adenines.

Q5:
Transcribing dna to rna
```{r}
my_rna <- my_dna %>% 
  str_replace_all("T", "U")
```
Finding how many start codons in the rna:
```{r}
sum(str_detect(my_rna, "AUG"))
```

One start codon in the rna.

Q6:
Extracting first three and last three nucleotides of my_dna:
```{r}
first_3_bases <- str_sub(my_dna, 1, 3)
last_3_bases <- str_sub(my_dna, -3, -1)
```

TCT encodes serine.

TGT encodes cysteine.

Q7:
Choosing AAT as random codon:
```{r}
rand_codon <- "AAT"
```
Now splitting my_dna on the random codon:
```{r}
str_split(my_dna, rand_codon)
```

4 substrings are returned.

Q8:
A factor adds a level of information which makes you able to sort strings on other levels than alphabetical.
For example, if we had a vector of the months of the year, it would not make sense to sort them alphabetically. However, we can make a new variable with the correct sorting of the months and use this variable to correctly sort vectors with months.

Q9:
```{r}
factor(LETTERS)
```
```{r}
factor(rep(LETTERS, 10))
```

This factor sorts alphabetically since no other factor level is given to the factor function.

Q10:
```{r}
factor(rev(LETTERS))
```

```{r}
factor(rev(LETTERS), levels = LETTERS)
```

The factor used is here still the alphabetically sorted alphabet, however the letters are reversed.

Q11:
The files are csv files with diabetes patient information.

Q12:
The function for reading csv files are called read_csv.

Q13:
Reading all files into separate variables and joining them together to recreate the diabetes data set:
```{r}
age <- read_csv("data/data/04_diabetes_id_age.csv")
chol <- read_csv("data/data/04_diabetes_id_chol.csv")
bp.1d <- read_csv("data/data/04_diabetes_id_bp.1d.csv")
bp.1s <- read_csv("data/data/04_diabetes_id_bp.1s.csv")
bp.2d <- read_csv("data/data/04_diabetes_id_bp.2d.csv")
bp.2s <- read_csv("data/data/04_diabetes_id_bp.2s.csv")
frame <- read_csv("data/data/04_diabetes_id_frame.csv")
gender <- read_csv("data/data/04_diabetes_id_gender.csv")
glyhb <- read_csv("data/data/04_diabetes_id_glyhb.csv")
hdl <- read_csv("data/data/04_diabetes_id_hdl.csv")
height <- read_csv("data/data/04_diabetes_id_height.csv")
hip <- read_csv("data/data/04_diabetes_id_hip.csv")
location <- read_csv("data/data/04_diabetes_id_location.csv")
ratio <- read_csv("data/data/04_diabetes_id_ratio.csv")
stab.glu <- read_csv("data/data/04_diabetes_id_stab.glu.csv")
time.ppn <- read_csv("data/data/04_diabetes_id_time.ppn.csv")
waist <- read_csv("data/data/04_diabetes_id_waist.csv")
weight <- read_csv("data/data/04_diabetes_id_weight.csv")
```

Joining the separate files together can be accomplished by full joining in a pipe:

```{r}
diabetes <- chol %>% 
  full_join(stab.glu, by = "id") %>% 
  full_join(hdl, by = "id") %>% 
  full_join(ratio, by = "id") %>%
  full_join(glyhb, by = "id") %>% 
  full_join(location, by = "id") %>% 
  full_join(age, by = "id") %>% 
  full_join(gender, by = "id") %>% 
  full_join(height, by = "id") %>% 
  full_join(weight, by = "id") %>% 
  full_join(frame, by = "id") %>% 
  full_join(bp.1s, by = "id") %>% 
  full_join(bp.1d, by = "id") %>% 
  full_join(bp.2s, by = "id") %>% 
  full_join(bp.2d, by = "id") %>%
  full_join(waist, by = "id") %>% 
  full_join(hip, by = "id")
  
```

Q14:
The diabetes data we have created is wide data.

Q15:
The function select_if selects column based on a predicate. If the predicate is TRUE, the column(s) will be selected.

Q16:
Converting from wide to long data with pivot_longer:
```{r}
diabetes_data_long <- select_if(diabetes, is.numeric) %>%
  pivot_longer(chol:hip, names_to = "attributes", values_to = "value")
  
```
Now plotting:
```{r}
ggplot(diabetes_data_long, mapping = aes(x = id, y = value)) +
  geom_point() +
  facet_wrap(~attributes)
```

Q17:
A histogram for each numerical variable would probably be a better way of presenting the data.

