---
title: "Session 2 - Data Visualisation using ggplot"
author: "Jeppe N Samuelsen"
date: "17 2 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We start by loading the needed libraries.

```{r}
library("tidyverse")
library("datamicroarray")
```

We will now ready and wrangle the data for visualisation.

```{r}
data('gravier', package = 'datamicroarray')
set.seed(676571)
cancer_data=mutate(as_tibble(pluck(gravier,"x")),y=pluck(gravier,"y"),pt_id=1:length(pluck(gravier, "y")),age=round(rnorm(length(pluck(gravier,"y")),mean=55,sd=10),1))
cancer_data=rename(cancer_data,event_label=y)
cancer_data$age_group=cut(cancer_data$age,breaks=seq(10,100,by=10))
```

Looking at the first 8 rows and first 8 columns of the data.

```{r}
select(slice(cancer_data,1:8),pt_id,age,age_group,event_label,1:5)
```

Q1: The data set is a cancer data set, which comes from a microarray which gene expressions for persons of different age.

Q2: We can see how many rows and columns which are in the data set by using nrow and ncol.
```{r}
nrow(cancer_data)
ncol(cancer_data)
```

Q3: The variables corresponds to each column and the observations corresponds to each row.

We will now plot 2 x 20 random normally distributed numbers.

```{r}
ggplot(data = tibble(x = rnorm(20), y = rnorm(20)),
       mapping = aes(x = x, y = y)) +
  geom_point()
```

T1:
Scatterplot of age against g2E09.
```{r}
ggplot(data = cancer_data) +
  geom_point(mapping = aes(x=age, y=g2E09))
```

T2:
Lineplot of age against g2E09.
```{r}
ggplot(data = cancer_data) +
  geom_line(mapping = aes(x=age, y=g2E09))
```

T3:
Boxplot of g2E09.
```{r}
ggplot(data = cancer_data, mapping = aes(x = "", y = g2E09)) + 
  geom_boxplot()
```

T4:
Histogram of g2E09.
```{r}
ggplot(data = cancer_data, mapping = aes(x = g2E09)) + 
  geom_histogram()
```

T5:
Densitogram of g2E09.
```{r}
ggplot(data = cancer_data, mapping = aes(x = g2E09)) + 
  geom_density()
```

Q4:
All geoms do not require both x and y. Fx boxplot and histogram only requires one variable.

T6:
Boxplot of g2E09 stratified on the variable event_label.
```{r}
ggplot(data = cancer_data, mapping = aes(x = "", y = g2E09)) + 
  geom_boxplot() +
  facet_grid(~event_label)
```

T7:
Densitogram of g2E09 stratifed on the variable event_label.
```{r}
ggplot(data = cancer_data, mapping = aes(x = g2E09)) + 
  geom_density() +
  facet_grid(~event_label)
```

T8:
Boxplot of g2E09 stratifed on the variable age_group.
```{r}
p <- ggplot(data = cancer_data, mapping = aes(x = age_group, y = g2E09)) + 
  geom_boxplot()
```

Then stratified on event_label with colors.
```{r}
p + geom_boxplot(aes(color = event_label)) + labs(x = "age group")
```

Now adding new x label.
```{r}
p + geom_boxplot(aes(color = event_label)) + labs(x = "age group")
```


T9:
Scatterplot of g2E09 versus age and stratifed on event_label. Also adding smoothing line and labels.
```{r}
p1 <- ggplot(data = cancer_data, mapping = aes(x = age, y = g2E09)) + 
  geom_point()
p1 + geom_point(aes(color = event_label)) + labs(x = "age") + geom_smooth(mapping = aes(x=age, y = g2E09))
```
T10:
Scatterplot of gene g2E09 versus gene g7F07 and stratifed on event_label. Also adding smoothing line and labels.
```{r}
p2 <- ggplot(data = cancer_data, mapping = aes(x = g2E09, y = g7F07)) + 
  geom_point()
p2 + geom_point(aes(color = event_label)) + geom_smooth(mapping = aes(x=g2E09, y = g7F07))
```
Then splitting into panes based on age group.
```{r}
p2 + geom_point(aes(color = event_label)) + geom_smooth(mapping = aes(x=g2E09, y = g7F07)) + facet_wrap(~ age_group)
```
Labeling titles.
```{r}
p2 + geom_point(aes(color = event_label)) + geom_smooth(mapping = aes(x=g2E09, y = g7F07)) + facet_wrap(~ age_group) + labs(color = "prognosis")
```
T11:
Recreating boxplot:
```{r}
p3 <- ggplot(data = cancer_data, mapping = aes(x = event_label, y = g1CNS507, fill=event_label)) + 
  geom_boxplot() +
  coord_flip() +
  labs(title = "A prognostic DNA signature for T1T2 node-negative breast cancer patients", subtitle = "Labelling: good = no event, poor = early metastasis", x = "Event after diagnosis", y = "Expression level of g1CNS507 (log2 transformed)")
p3 + labs(fill = "")
```

Q5:
It is visble that a higher level of the gene g1CNS507 gives a better chance of a good event after diagnosis.

T12:
Recreating plot:
```{r}
p4 <- ggplot(data = cancer_data, mapping = aes(x = age, y = g1int239, color = event_label)) + 
  geom_point() +
  geom_smooth(mapping = aes(x=age, y = g1int239, color = event_label), method = "lm", se = FALSE)
p4 + labs(title = "A prognostic DNA signature for T1T2 node-negative breast cancer patients", subtitle = "Stratified on labelling: good = no event, poor = early metastasis", x = "Subject age", y = "Expression level of g1int239 (log2 transformed)", color = "Label")
```

