---
title: "Session 4 - Modelling, dimension reduction and clustering"
author: "Jeppe N Samuelsen"
date: "7 3 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

T1:

```{r}
library("tidyverse")
library("modelr")
options(na.action = na.warn)
diabetes_data <- read_csv(file = "data/diabetes.csv")
```

Convering height and weight to metric system:
```{r}
diabetes_data <- diabetes_data %>%
  mutate(height = 2.54*height) %>%
  mutate(height = round(height, 1)) %>%
  mutate(weight = 0.454*weight) %>%
  mutate(weight = round(weight, 1))
```


Recreating plot:
```{r}
ggplot(diabetes_data, mapping = aes(x = height, y = weight)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(gender ~ location)
```

T2:
Creating nested tibble:
```{r}
by_location <- diabetes_data %>%
  group_by(location, gender) %>%
  nest()
```

Creating weight model function:
```{r}
weight_model <- function(df) {
  lm(weight ~ height, data = df)
}
```

Defining models:
```{r}
models <- map(by_location$data, weight_model)
```

Applying the linear models to the all observations of the data frame:
```{r}
by_location <- by_location %>% 
  mutate(model = map(data, weight_model))
```

T3:
```

Unnesting the tibble with broom::glance:
```{r}
by_location %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)
```

T4:
Recreating tibble with broom::tidy:
```{r}
by_location_tidy <- by_location %>%  
  mutate(tidy = map(model, broom::tidy)) %>% 
  unnest(tidy)
```

T5:
```{r}
broom::confint_tidy(x = by_location_tidy, conf.level = 0.95)
```

T6:
Recreating plot:

