---
title: "Session 3 - Data manipulation using dplyr"
author: "Jeppe N Samuelsen"
date: "17 2 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1:
Shows the numbers that are above in a tibble.

Q2:
Orders the numbers in x by ascending order.

Q3:
Orders the numbers in x by descending order.

Q4:
Orders the numbers in x by ascending order.

Q5:
Selects only the x column (variable).

Q6:
Selects only the y column (variable).

Q7:
Selects only the x column (variable) multiplied by -1.

Q8:
Selects the x column (variable) multiplied by -1 and the y column (variable) multiplied by -1.

Q9:
Adds a column with the variable x_dbl = 2*x.

Q10:
Adds a column with the variable x_dbl = 2 * x, and a column with the variable x_qdr = 2 * x_dbl.

Q11:
Summarizes the mean of x values in the column x_mu.

Q12:
Summarizes the max value of x values in the variable x_max.

Q13:
First groups the data by the lbl variable, then summarizes the mean of x (x_mu) and max of x (x_max).

Q14:
First groups the data by the lbl variable, then summarizes the number of counts (observations) in the tubble as the variable n.

Q15:
First groups the data by the lbl variable, then summarizes the number of counts (observations) in the lbl variable/column.

```{r}
install.packages("tidyverse")
```


Reading diabetes data file:
```{r}
library("tidyverse")
diabetes_data <- read_csv(file = "data/diabetes.csv")
```

```{r}
diabetes_data
```

Q1:
403 observations and 19 variables

Q2:
It looks like a tidy dataset.
The rules that must be satisfied are:
1. Each variable must have its own column.
2. Each observation must have its own row.
3. Each value must have its own cell.

Q3:
The types of the different variables.

T1:
Changing variables height, weight, waist and hip from inches/pounds to cm/kg, rounding to 1 decimal
```{r}
diabetes_data <- diabetes_data %>% mutate(height = 2.54*height) %>% mutate(height = round(height, 1)) %>% mutate(weight = 0.454*weight) %>% mutate(weight = round(weight, 1)) %>% 
mutate(waist = 2.54*waist) %>% mutate(waist = round(waist, 1)) %>% 
mutate(hip = 2.54*hip) %>% mutate(hip = round(hip, 1))
```

Q4:
```{r}
diabetes_data %>% filter(weight < 100)
```
349 persons weigh less than 100 kg.

Q5:
```{r}
diabetes_data %>% filter(weight > 100)
```

53 persons weigh more than 100 kg.

Q6:
```{r}
diabetes_data %>% filter(weight > 100, height < 160)
```
2 persons weigh more than 100 kg and are less than 1.6 m in height.

Q7:
```{r}
diabetes_data %>% filter(gender == "female", height > 180)
```

4 women are taller than 1.8m.

Q8:
```{r}
diabetes_data %>% filter(gender == "male", height > 180)
```

51 men are taller than 1.8m.

Q9:
```{r}
diabetes_data %>% filter(location == "Louisa", age > 30, gender == "female")
```

95 women from Louisa are older than 30.

Q10:
```{r}
diabetes_data %>% filter(gender == "male", location == "Buckingham", height > 190)
```

5 males from Buckingham are taller than 1.9m.

T2:
Scatter plot of weight versus height and colored by gender for inhabitants of Louisa above the age of 40.
```{r}
scatter1 <- diabetes_data %>% filter(location == "Louisa", age > 40)
ggplot(data = scatter1, mapping = aes(x = weight, y = height)) + 
  geom_point() +
  geom_point(aes(color = gender))
```

T3:
Boxplot of height versus location stratified on gender for people above the age of 50.
```{r}
Box1 <- diabetes_data %>% filter(age > 50)
ggplot(data = Box1, mapping = aes(x = location, y = height)) + 
  geom_boxplot() +
  geom_boxplot(aes(color = gender))
```

Q11:
```{r}
diabetes_data %>% summarise(min(age))
```

The youngest person is 19 years old.

Q12:
```{r}
diabetes_data %>% summarise(max(age))
```

The oldest person is 92 years old.

Q13:
```{r}
diabetes_data %>% filter(age == 20) %>% 
  summarise(max(height))
```

Of the 20 years olds, the tallest person is 182.9 cm.

Q14:
```{r}
diabetes_data %>% filter(age == 20) %>% summarise(min(height))
```

Of the 20 years olds, the shortest person is 147.3 cm.

Q15:
```{r}
diabetes_data %>% 
  select(starts_with("b"))
```

4 variables starts wtih a b.

Q16:
```{r}
diabetes_data %>% 
  select(contains("eight"))
```

2 variables contains the word eight.

T4:
Creating new variable where BMI is calculated:
```{r}
diabetes_data <- diabetes_data %>% mutate(BMI = weight/((height/100)^2))
diabetes_data
```

T5:
Creating a BMI_class variable:
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(BMI_class = case_when(BMI < 18.5 ~ "underweight", BMI >= 18.5 & BMI <= 24.9 ~ "normal weight", BMI >= 25.0 & BMI <= 29.9 ~ "overweight", BMI >= 30.0 & BMI <= 34.9 ~ "obese", BMI >= 35.0 & BMI <= 39.9 ~ "severe obesity", BMI >= 40 ~ "morbid obesity"))
```

Then making BMI_class a categorical class by "factor":
```{r}
diabetes_data <- diabetes_data %>%
  mutate(BMI_class = factor(BMI_class,
                            levels =  c("underweight", "normal weight", "overweight", "obese",
                                        "severe obesity", "morbid obesity")))
```

T6:
Making boxplot of hdl versus BMI_class:
```{r}
ggplot(data = diabetes_data, mapping = aes(x = BMI_class, y = hdl)) +
  geom_boxplot()
```

Q17:
High values of hdl is generally observed in underweight persons.

T8:
First, i create a variable called gender_discrete, where 0 corresponds to female and 1 corresponds to male.

```{r}
diabetes_data <- diabetes_data %>% 
  mutate(gender_discrete = case_when(gender == "female" ~ 0, gender == "male" ~ 1))
```

(Deleting fail column):

```{r}
diabetes_data <- diabetes_data %>% select(-c(age_discrete))
```

Now we make a new BFP column where the BFP is calculated:
```{r}
diabetes_data <- diabetes_data %>%
  mutate(BFP = 1.39*BMI+0.16*age-10.34*gender_discrete-9)
```

T8:
Creating a waist-to-hip ratio:
```{r}
diabetes_data <- diabetes_data %>% 
  mutate(WHR = waist/hip)
```

Q18:
Scatterplot of BFP versus BMI:
```{r}
ggplot(data = diabetes_data, aes(x = BFP, y = BMI)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Now, creating scatterplot of WHR versus BMI:
```{r}
ggplot(data = diabetes_data, aes(x = WHR, y = BMI)) +
  geom_point() +
  geom_smooth(method = "lm")
```

We can clearly see from the plots that body fat percentage (BFP) correlates well with BMI, while waist-to-hip ratio doesn't.

Q19:
```{r}
diabetes_data %>%
  filter(gender == "male")
```

```{r}
diabetes_data %>%
  filter(gender == "female")
```

169 males and 234 females in the dataset.

Q20:
```{r}
diabetes_data %>% 
  filter(gender == "male", location == "Buckingham")
```

```{r}
diabetes_data %>% 
  filter(gender == "male", location == "Louisa")
```

86 males from Buckingham and 83 males from Louisa.

```{r}
diabetes_data %>% 
  filter(gender == "female", location == "Buckingham")
```

```{r}
diabetes_data %>% 
  filter(gender == "female", location == "Louisa")
```

114 females from Buckingham and 120 females from Louisa.

Q21:

