---
title: "Session 1 - R Basics"
author: "Jeppe N Samuelsen"
date: "7 2 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this section I will simulate a dice.

I will start by defining a die that goes from 1 to 6.
```{r}
die <- 1:6

die
```

We can roll the die with the function "sample" that takes two arguments:
a vector x and a size.

I will now roll the die two times.

```{r}
sample(x=die, size=2)
```

However that was without replacement so we can not get the same number in the two rolls.

We can also roll the die with replacement:

```{r}
sample(x=die, size=2, replace=TRUE)
```

We can also add up the two rolls.

```{r}
dice <- sample(x=die, size=2, replace=TRUE)

sum(dice)
```

The dice will give the same result every time (the result that was rolled when the dice was defined)

```{r}
dice

dice

dice
```

We will now create a function called roll(), which can roll new numbers every time it is called.

Functions consist of a name, a body of code, and a set of arguments. The output of the function is the last line of the body of code (if the line has an output).

We could make the roll function without any arguments, like this:

```{r}
roll <- function() {
  die <- 1:6
  dice <- sample(x=die, size=2, replace=TRUE)
  sum(dice)
}
```

We can now run this roll function.

```{r}
roll()
```

Typing the function name without the parenthesis will not run the function but just give the code of the function.

```{r}
roll

roll()
```

We can also give our function arguments (think of them like input to the function). Like this:

```{r}
roll2 <- function(bones){
  dice <- sample(x=bones, size=2, replace = TRUE)
  sum(dice)
}
```

When running the function we now need to input the argument "bones", which should be a vector to use as input in the sample function.

```{r}
roll2(bones = 1:4)
roll2(bones = 1:6)
roll2(1:20)
```
Now the function roll2 needs a supply of the bones vector, otherwise it will give an error message.

We can prevent this error by supplying the function with a default bones value.

```{r}
roll2 <- function(bones = 1:6){
  dice <- sample(x = bones, size = 2, replace = TRUE)
  sum(dice)
}

roll2()
```

We have now created our own dice.
